Split-Apply-Combine with Dynamic Grouping

### 2023-03-14 paper/article.Rnw
wip

\subsection{Imputing missing values using SAE and ratio imputation} \label{sect:ratio}
Our goal in this example is to impute missing values for the \code{industrial}
variable based on ratio imputation with \code{total} as predictor.  Ratio
imputation is a method where the imputed value $\hat{y}_i$ for variable $Y$ of
record $i$ is estimated as $\hat{y_i}=\hat{R}_d x_i$, where $\hat{R}$ is an
estimate of the ratio between $Y$ and an auxiliary variable $X$ in
subpopulation $d$. An unbiased estimate for $R_d$ is given by
$\hat{\bar{Y}}_d/\hat{\bar{X}}_d$ where $\hat{\bar{Y}}_d$ and $\hat{\bar{X}}_d$
are estimated subpopulation means.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Random Nearest neighbours imputation with collapsing groups}
Nearest neighbor (NN) imputation is a donor imputation method where the
imputation value is copied from a (possibly random) donor record with
attributes that are sufficiently similar \citep{andridge2010review}.
One way to generate an imputation value is to select a random donor
from the same group. To prevent the same donor from being used too
often, it is not uncommon to demand a minimum number of records
in the donor pool. A collapsing scheme is one way of guaranteeing
this, and below we demonstrate how this problem can be expressed
in \code{accumulate}.

@@

We first define an `aggregation' function that takes a vector 
of donors and returns a non-empty sample.
<<>>=
random_element <- function(x) sample(x[!is.na(x)], 1)
@
To make sure we obtain a donor for each record, we add an
identifying column to use as grouping variable.
\code{id} column.
<<>>=
producers <- cbind(id = sprintf("ID%03d",seq_len(nrow(producers)))
                  , producers)
set.seed(111)
imputations <- cumulate(producers
                 , collapse = id ~ sbi*size + sbi + sbi2 + sbi1
                 , test = min_complete(5,"trade")
                 , donor_trade = random_element(trade))

>>>>

\subsection{Imputing missing values using SAE and ratio imputation} \label{sect:ratio}
Our goal in this example is to impute missing values for the \code{industrial}
variable based on ratio imputation with \code{total} as predictor.  Ratio
imputation is a method where the imputed value $\hat{y}_i$ for variable $Y$ of
record $i$ is estimated as $\hat{y_i}=\hat{R}_d x_i$, where $\hat{R}_d$ is an
estimate of the ratio between $Y$ and an auxiliary variable $X$ in
subpopulation $d$. An unbiased estimate for $R_d$ is given by
$\hat{\bar{Y}}_d/\hat{\bar{X}}_d$ where $\hat{\bar{Y}}_d$ and $\hat{\bar{X}}_d$
are estimated subpopulation means.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Random Nearest neighbours imputation with collapsing groups}
Nearest neighbor (NN) imputation is a donor imputation method where the
imputation value is copied from a record that is (randomly) chosen from a donor
pool \citep{andridge2010review}. In this example we use the grouping variables
in \code{producers} to define the donor pools. To prevent the same donor from
being used too often, it is not uncommon to demand a minimum number of records
in the donor pool. A collapsing scheme is one way of guaranteeing this, and
below we demonstrate how this problem can be expressed in \code{accumulate}.

@@

We first define an `aggregation' function that takes a vector 
of donors and returns a non-empty sample.
<<>>=
random_element <- function(x) sample(x[!is.na(x)], 1)
@
We will use \code{cumulate()} to ensure that there are at least five non-empty
values in \code{x} when \code{random_element()} is called.

To make sure we obtain a donor for each record, we add an
identifying column to use as grouping variable.
\code{id} column.
<<>>=
producers <- cbind(id = sprintf("ID%03d",seq_len(nrow(producers)))
                  , producers)
set.seed(111)
imputations <- cumulate(producers
                 , collapse = id ~ sbi*size + sbi + sbi2 + sbi1
                 , test = min_complete(5,"trade")
                 , donor_trade = random_element(trade))
